---
title: "details_and_wardsAnalysisG6"
author: "Hannah Brown"
date: "1/7/2020"
output: html_document
---

```{r}
#loading data
crashes <- read_csv("https://opendata.arcgis.com/datasets/70392a096a8e431381f1f692aaa06afd_24.csv")
  
crashes$WARD[crashes$WARD == "Null"] <- NA

summary(is.na(crashes))
#crashes2 <- na.omit(crashes)

# throw out columns
crashes$LOCATIONERROR <- crashes$LASTUPDATEDATE <- crashes$MPDLATITUDE <- crashes$MPDLONGITUDE <- crashes$MPDGEOX <- crashes$MPDGEOY <- crashes$STREETSEGID <- crashes$ROADWAYSEGID <- crashes$MAR_ADDRESS <- crashes$TODATE <- crashes$FATALPASSENGER <- crashes$MAJORINJURIESPASSENGER <- crashes$UNKNOWNINJURIESPASSENGER <- NULL
crashes <- crashes %>% filter(as.Date(FROMDATE) >= "2012-01-01")

crashes <- crashes %>%
  mutate(INDEX = 1)
View(crashes)

details_and_wards <- read.csv("https://opendata.arcgis.com/datasets/70248b73c20f46b0a5ee895fc91d6222_25.csv")
```

```{r}
library(tidyverse)
library(magrittr)
library(ggplot2)
```


```{r}
#summarizing data
sumDet <- summary(details_and_wards)
sumDet
```

```{r}
# details_and_wards$WARD <- crashes$WARD[match(crashes$CRIMEID,crashes$CRIMEID), all = FALSE];
details_and_wards <- left_join(details_and_wards, crashes, by = "CRIMEID")
```


```{r}
#table summaries
#yikes forgot this was continuous lol table(details_and_wards$AGE)
table(details_and_wards$INVEHICLETYPE)
table(details_and_wards$LICENSEPLATESTATE)
```

```{r}
#filtering out ages that are not relevant
details_and_wards %>% 
  filter((AGE >= 16)&(AGE <= 93)) ->
  age1
summary(age1)
```

```{r}
details_and_wards %>% 
  mutate(`FATAL` = recode(FATAL, "N" = 0, 
                           "Y" = 1 ),
          `MAJORINJURY` = recode(`MAJORINJURY`, "N" = 0, 
                                    "Y" = 1 ),
         `MINORINJURY` = recode(MINORINJURY,"N" = 0, 
                                       "Y" = 1 ),
         `TICKETISSUED` = recode(TICKETISSUED,"N" = 0, 
                                       "Y" = 1 ),
         `IMPAIRED` = recode(IMPAIRED,"N" = 0, 
                                       "Y" = 1 ),
         `SPEEDING` = recode(SPEEDING,"N" = 0, 
                                       "Y" = 1 )) ->
  details_and_wards1
```

```{r}
#filtering out ages that are not relevant
details_and_wards1 %>% 
  filter((AGE >= 16)&(AGE <= 93)) ->
  details_and_wards1
```

```{r}
#filtering out ages that are not relevant
details_and_wards1 %>% 
  filter((LICENSEPLATESTATE == "MD")|(LICENSEPLATESTATE == "VA")|(LICENSEPLATESTATE == "DC")) ->
  details_and_wards1
summary(details_and_wards1)
```

```{r}
#plotting relevant ages
ggplot(details_and_wards1 = details_and_wards1, mapping = aes(x = details_and_wards1$AGE, fill = details_and_wards1$LICENSEPLATESTATE)) +
  geom_histogram( bins = 20)
```

```{r}
details_and_wards1 %>% 
  ggplot(data = details_and_wards1, mapping = aes(x = LICENSEPLATESTATE)) +
  geom_bar()
```

```{r}
details_and_wards1 %>% 
  ggplot(data = details_and_wards1, mapping = aes(INVEHICLETYPE, fill = FATAL), par(las = 2)) +
  geom_bar(position = "stack")
```

```{r}
details_and_wards %>% 
  ggplot(data = details_and_wards, mapping = aes(PERSONTYPE)) +
  geom_bar(position = "stack", las = 2)
```


```{r}
details_and_wards1 %>% 
  ggplot(data = details_and_wards1, mapping = aes(PERSONTYPE, FATAL, color = LICENSEPLATESTATE)) +
  geom_jitter() +
  facet_wrap(~INVEHICLETYPE)
```


```{r}
details_and_wards1 %>% 
  ggplot(data = details_and_wards1, mapping = aes(x = MAJORINJURY,  y = AGE, color = IMPAIRED)) +
  geom_bar(stat = "identity")
```



#Set Up
```{r}
library(readr)
library(dplyr)
library(here)
library(MCMCpack)
library(glmnet)
library(parsnip)
```

#Estimate Bayesian Logistic Regression

Est of Bayesian logistic regression
```{r}
mc_post <- MCMClogit(FATAL ~ 1, data = details_and_wards1)
summary(mc_post)
```

Modeling fuller Bayesian model
```{r}
mc_post_full <- MCMClogit(FATAL ~ MAJORINJURY + MINORINJURY + TICKETISSUED + IMPAIRED + SPEEDING + AGE, data = details_and_wards1, burnin = 500, mcmc = 2000, thin = 2)
summary(mc_post_full)
```

Highest posterior density intervals for the coef
```{r}
HPDinterval(mc_post_full) -> mc_int
mc_int
```

```{r}
pdf("mcmc_details_and_wards.pdf")
plot(mc_post_full)
dev.off()
```

Posterior Probability Calculation- What is the posterior probability the `Neighbors` postcard has a larger coefficient than the `Hawthorne` card?
```{r}
mean(mc_post_full[, "MINORINJURY"] < mc_post_full[, "MAJORINJURY"])
```

Estimate Machine Learning Models

`X` is a raw numeric matrix; `y` is a raw numeric vector
```{r}
predictors <- c("MAJORINJURY", "MINORINJURY", "TICKETISSUED", "IMPAIRED", "SPEEDING", "AGE")

X <- details_and_wards1[, predictors]
```

Feature engineering
```{r}
X <- X %>% mutate(age2 = AGE^2,
                  age3 = AGE^3,
                  age4 = AGE^4,
                  age5 = AGE^5) %>%
  as.matrix()

y <- details_and_wards1[, "FATAL"] %>% unlist %>% as.numeric()
```

Est LS model
```{r}
lm_out <- lm(FATAL ~ MAJORINJURY + MINORINJURY + TICKETISSUED + IMPAIRED + SPEEDING + AGE + 
            I(AGE^2) + I(AGE^3) + I(AGE^4) + I(AGE^5),
             data = details_and_wards1)
coefs_lm <- coef(lm_out)
summary(lm_out)
```

Estimate with LASSO
```{r}
lasso_out <- glmnet(X, y, alpha = 1)
lasso_out
```

How to choose coefs? Use cross-validation to see which predicts best in `out-of-sample` tests. 
```{r}
set.seed(281) #train-test sets are consistent across implementations
cv_lasso_out <- cv.glmnet(X, y, alpha = 1)
```

Show best coefs for the model, the one with the min deviance
```{r}
coef(cv_lasso_out, s = "lambda.min")
```

Lamda within in 1 SE of the min
```{r}
coef(cv_lasso_out, s = "lambda.1se")
```

```{r}
coefs_lasso <- coef(cv_lasso_out, s = "lambda.1se")
```

```{r}
cbind(coefs_lasso, coefs_lm)
```



