---
title: "map_slides"
author: "Lucas Gorak"
date: "1/10/2020"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(knitr)
library(lubridate)
library(ggplot2)
library(dplyr)
library(data.table)
library(ggrepel)
library(tidyverse)
library("ggmap")
library(broom)


col1 = "#011f4b"
col2 = "#6497b1"
col3 = "#b3cde0"
col4 = "#CC0000"
```

##Required Libraries

```{r}
# library(lubridate)
# library(ggplot2)
# library(dplyr)
# library(data.table)
# library(ggrepel)
# library(tidyverse)
# library("ggmap")
# library(broom)
```

##Threeoneone Data Import

```{r}
threeoneone_2 = read_csv("https://datagate.dc.gov/search/open/311requests?daterange=8years&details=true&format=csv")

head(threeoneone_2)
attach(threeoneone_2)

threeoneone_2long <- threeoneone_2$LONGITUDE
threeoneone_2lat <- threeoneone_2$LATITUDE
```

##Crash Data Import

```{r}
crashdata <- read.csv("https://opendata.arcgis.com/datasets/70392a096a8e431381f1f692aaa06afd_24.csv")

head(crashdata)
attach(crashdata)
```

##Data Wrangling

```{r}
threeoneone_2$ymd <- ymd_hms(ADDDATE)

threeoneone_2$ymd 

threeoneone_2[complete.cases(threeoneone_2), ]

crashdata_clean_date <- gsub("T.*", "", crashdata$FROMDATE)

crashdata$crashdate <- ymd(crashdata_clean_date)

crashdata[complete.cases(crashdata), ]

crashdatalong <- crashdata$LONGITUDE
crashdatalat <- crashdata$LATITUDE
```

##Create DC Map with Google API

```{r}
ggmap::register_google(key = "AIzaSyA3nim7i8YYK3_Bkz9oqFW1Vcg0rXDG7lY")


dcmap <- ggmap(get_map(location = c(lon = -77.02106, lat = 38.905), zoom = 12, scale = 1, maptype = 'terrain', color = 'color'))
```

##DC Map 

```{r}
dcmap
```

##Threeoneone Requests in DC

```{r}
dcmap_311 <- dcmap + geom_point(data = threeoneone_2, aes(x = threeoneone_2long, y = threeoneone_2lat), color = col1, alpha = 0.05, size = 0.05) + theme(legend.position = "none") 

dcmap_311
```




##Crashes in DC

```{r}
dcmap_crashes <- dcmap + geom_point(data = crashdata, aes(x = crashdatalong, y = crashdatalat), color = col4, alpha = 0.05, size = 0.05) + theme(legend.position = "none")

dcmap_crashes
```

##Threeoneone and Crashes in DC

```{r}
dcmap_311_crashes <- dcmap + geom_point(data = threeoneone_2, aes(x = threeoneone_2long, y = threeoneone_2lat), color = col1, alpha = 0.05, size = 0.05) + geom_point(data = crashdata, aes(x = crashdatalong, y = crashdatalat), color = col4, alpha = 0.05, size = 0.05) + theme(legend.position = "none") 

dcmap_311_crashes
```

##Detail of Threeoneone and Crashes in Downtown DC

```{r}
dcmap_dt <- ggmap(get_map(location = c(lon = -77.02906, lat = 38.905), zoom = 14, scale = 2, maptype = 'terrain', color = 'color'))

dcmap_dt_311_crashes <- dcmap_dt + geom_point(data = threeoneone_2, aes(x = threeoneone_2long, y = threeoneone_2lat), color = col1, alpha = 0.05, size = 1) + geom_point(data = crashdata, aes(x = crashdatalong, y = crashdatalat), color = col4, alpha = 0.05, size = 1) + theme(legend.position = "none")

dcmap_dt_311_crashes
```

##Detail of Threeoneone and Crashes in NW DC

```{r}
dcmap_nw <- ggmap(get_map(location = c(lon = -77.07, lat = 38.94), zoom = 14, scale = 2, maptype = 'terrain', color = 'color'))

dcmap_nw_311_crashes <- dcmap_nw + geom_point(data = threeoneone_2, aes(x = threeoneone_2long, y = threeoneone_2lat), color = col1, alpha = 0.05, size = 1) + geom_point(data = crashdata, aes(x = crashdatalong, y = crashdatalat), color = col4, alpha = 0.05, size = 1) + theme(legend.position = "none")

dcmap_nw_311_crashes
```

##Detail of Threeoneone and Crashes in NE DC

```{r}
dcmap_ne <- ggmap(get_map(location = c(lon = -76.995, lat = 38.925), zoom = 14, scale = 2, maptype = 'terrain', color = 'color'))

dcmap_ne_311_crashes <- dcmap_ne + geom_point(data = threeoneone_2, aes(x = threeoneone_2long, y = threeoneone_2lat), color = col1, alpha = 0.05, size = 1) + geom_point(data = crashdata, aes(x = crashdatalong, y = crashdatalat), color = col4, alpha = 0.05, size = 1) + theme(legend.position = "none")

dcmap_ne_311_crashes
```

##Detail of Threeoneone and Crashes in SW/SE DC

```{r}
dcmap_swse <- ggmap(get_map(location = c(lon = -76.993, lat = 38.867), zoom = 14, scale = 2, maptype = 'terrain', color = 'color'))

dcmap_swse_311_crashes <- dcmap_swse + geom_point(data = threeoneone_2, aes(x = threeoneone_2long, y = threeoneone_2lat), color = col1, alpha = 0.05, size = 1) + geom_point(data = crashdata, aes(x = crashdatalong, y = crashdatalat), color = col4, alpha = 0.05, size = 1) + theme(legend.position = "none")

dcmap_swse_311_crashes
```

##Clustering Crash Coordinates (Pre-Cluster Wrangling)

```{r}
crashdata <- crashdata %>% filter( X > -77.11, X < -76.89)
crashdata <- crashdata %>% filter( Y < 39, Y > 38.7)

crashx <- crashdata$X
crashy <- crashdata$Y 

crashxy <- matrix(c(crashx,crashy), ncol = 2)
```

#Making Two Sets of Clusters

```{r}
crash_clusterinfo_8 <- kmeans(crashxy, 8)
crash_clusterinfo_15 <- kmeans(crashxy, 15)

crashxy_8 <- augment(crash_clusterinfo_8, crashxy)
crashxy_15 <- augment(crash_clusterinfo_15, crashxy)
```

#Mapping the Clusters

```{r}
dcmap_crash_kcluster_8 <- dcmap + geom_point(data = crashxy_8, aes(x = crashxy_8$X1, y = crashxy_8$X2, color = crashxy_8$.cluster, alpha = 0.01, size = 0.001)) + theme(legend.position = "none")

dcmap_crash_kcluster_15 <- dcmap + geom_point(data = crashxy_15, aes(x = crashxy_15$X1, y = crashxy_15$X2, color = crashxy_15$.cluster, alpha = 0.01, size = 0.001)) + theme(legend.position = "none")

dcmap_crash_kcluster_8

dcmap_crash_kcluster_15

```

##Clustering Threeoneone Coordinates (Pre-Cluster Wrangling)

```{r}
threeoneonex <- threeoneone_2$LATITUDE
threeoneoney <- threeoneone_2$LONGITUDE

threeoneonexy <- matrix(c(threeoneonex,threeoneoney), ncol = 2)
```

##Making Two Sets of Clusters

```{r}
threeoneone_cluster_8 <- kmeans(threeoneonexy, 8)
threeoneone_cluster_15 <- kmeans(threeoneonexy, 15)

threeoneonexy_8 <- augment(threeoneone_cluster_8, threeoneonexy)
threeoneonexy_15 <- augment(threeoneone_cluster_15, threeoneonexy)
```

## Mapping the Clusters

```{r}
dcmap_threeoneone_kcluster_8 <- dcmap + geom_point(data = threeoneonexy_8, aes(x = threeoneonexy_8$X1, y = threeoneonexy_8$X2, color = threeoneonexy_8$.cluster, alpha = 0.01, size = 0.001)) + theme(legend.position = "none")

dcmap_threeoneone_kcluster_15 <- dcmap + geom_point(data = threeoneonexy_15, aes(x = threeoneonexy_15$X1, y = threeoneonexy_15$X2, colour = threeoneonexy_15$.cluster, alpha = 0.01, size = 0.001)) + theme(legend.position = "none")

dcmap_threeoneone_kcluster_8

dcmap_threeoneone_kcluster_15
```

##Comparing the Two Sets of Clusters

```{r}
dcmap_crash_kcluster_8

dcmap_threeoneone_kcluster_8
```

## 

```{r}
dcmap_crash_kcluster_15

dcmap_threeoneone_kcluster_15
```

